---
title: "Manuscript for MADA Project"
subtitle: ""
author: Jacob Matta
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../dataanalysis_template_references.bib
csl: ../apa.csl
---
# Part 1

## Description of data and data source

The data is a count of residents with at least one COVID-19 vaccination by demographic in the state of Georgia released by the Georgia Department of Public Health. The data has 1962 observations. The data is uploaded weekly.

You can download the data from this link: <https://www.arcgis.com/home/item.html?id=17a8a4af79e5453a9132aaa29fb4ad22>

## Why did I choose this data?

As a graduate research assistant, I am participating in an evaluation being conducted on a COVID-19 health literacy and vaccine hesitancy program in Albany, GA. This secondary data will be used to determine whether the program will produce a significant increase in COVID-19 vaccinations within the target zipcodes of 31701, 31705 and 31707.

##Loading the Data 

```{r}
#looading packages
library(readxl) #for loading Excel files
library(here) #to set paths
library(dplyr) #for description summary of data 
```

```{r}
#loading data
rawdata_location <- here::here("data","raw_data","rawdata.xlsx")
rawdata <- readxl::read_excel(rawdata_location)

#loading data code book for reference 
codebook_location <-here::here("data","raw_data","Code_book_Vaccination_Demog_.xlsx")

codebook <- readxl::read_excel(codebook_location)
```

##Summary Descriptions of Data
```{r}
summary(rawdata)
```

## Questions/Hypotheses to be addressed
Using this data, I would like to explore whether the COVID-19 health literacy and vaccine hesitancy program in Albany, GA is causing a noticeable increase in vaccinations within the zipcodes 31701, 31705, 31707. 

##Predictors and Outcomes
My goal is to stratify the data by race and age and observe the difference due to those demographics. 

##Overall Goal 
Since the data is updated weekly by the Georgia Department of Public Health, I am seeking to establish an efficient method of processing and cleaning the data to produce a data set with the target zip codes. Then I intend to analyze in hopes of observing the amount of vaccinations in the given week. I can then apply that data cleaning and analyzing process to each proceeding week to compare results and examine whether there is an increase in vaccination rates within the time of the health literacy program. 

##Important Notes
I have been downloading this data weekly since September 17th and I will continue doing so until the end of the intervention in Albany. 


