---
title: "An example cleaning script"
author: "Andreas Handel"
date: "2023-01-03"
output: html_document
---

# Processing script

This contains the same code and comments/information as `processingcode.R`.

This just shows it as Quarto file, to give you an idea how to do it in a setup that combines code and text in a single file.

# Setup

Load needed packages. make sure they are installed.

```{r}
library(readxl) #for loading Excel files
library(here) #used for setting a path 
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
```

# Raw Data loading

```{r}
rawdata_location = here("data","raw_data","Sept_17.xlsx")
Sept17 = read_excel(rawdata_location)
glimpse(Sept17)
```

# Codebook loading

```{r}
#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

# Summary of raw data

```{r}
glimpse(Sept_17)
summary(Sept_17)
```

# Filtering the data by Doughert County
### I am only interested in Covid-19 Vaccination from Dougherty County which contains the target zip codes of 31701, 31705 and 31707
```{r}
Dougherty_Vax_Week1 <- filter(Sept_17, COUNTY_NAME == "Dougherty County")  

#replacing negative values with 0
Dougherty_Vax_Week1 [Dougherty_Vax_Week1 < 0] <- 0  
```
